---
title: "The exploration's report"
date: "August, 2019"
---

```{r setup_exploration, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos =  "h")
knitr::opts_knit$set(root.dir = "../")

# loading required libraries
library(rmarkdown)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(ggalluvial)
library(tidyr)
library(lubridate)
library(stringr)
library(VIM)
library(psych)
library(feather)
library(tinytex)
library(knitr)
library(leaflet)
library(geojsonio)
```

```{r scripts_exploration, include=FALSE}
# loading required steps before performing the analysis
source("./scripts/step_01_create_functions.R")
source("./scripts/step_02_data_ingestion.R")
source("./scripts/step_03_data_cleaning.R")
source("./scripts/step_04_label_translation.R")
source("./scripts/step_05_data_enhancement.R")
```

# Gender Exploration

At first glance, gender equality is well balanced in the bank, even when observed over the decades. Even more impressive, gender equality is everywhere in the country.

```{r gender_distribution_analysis, echo=TRUE, out.width = '100%'}
# gender distribution of clients in the bank
gender_plot <- ggplot(data = client) +
  aes(x = gender, fill = gender) +
  geom_bar() +
  labs(subtitle = "A well balanced bank",
       x = "Gender",
       y = "Total clients") +
  theme_economist()

clientGenderOverDecades <- client %>% 
  group_by(decade = as.integer(substr(client$birth_number, 1,1)) * 10, 
           gender = client$gender) %>% 
  count()

# gender distribution of clients in the bank over the decades
gender_dist_plot <- ggplot(data = clientGenderOverDecades) +
  aes(x = decade, fill = gender, weight = n) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
  geom_bar() +
  geom_line(aes(y = n, color = gender)) +
  labs(subtitle = "Equality at its finest",
       x = "Decades",
       y = "Total clients") +
  theme_economist() +
  facet_wrap(vars(gender))

gender_plots <- ggarrange(gender_plot, gender_dist_plot)

gender_plots <- annotate_figure(gender_plots, 
                                 top = text_grob("Gender distribution of clients over the decades", 
                                                 color = "black", face = "bold", 
                                                 size = 14))
gender_plots

# alluvial diagram representation of gender, age group and region
clientGenderAgeGroupByRegion <- client %>% 
  mutate(age_group = ifelse(age < 21, "young", 
                     ifelse(age >= 21 & age <= 60, "adult", "senior"))) %>% 
  inner_join(district, by = "district_id") %>% 
  group_by(age_group, gender, region) %>% 
  count()
  
ggplot(data = clientGenderAgeGroupByRegion, 
       aes(axis1 = region, axis2 = age_group, y = n)) +
  scale_x_discrete(limits = c("region", "age group"), expand = c(.1, .1)) +
  xlab("Demographic") +
  geom_alluvium(aes(fill = gender), knot.pos = 0) +
  geom_stratum() + 
  geom_text(stat = "stratum", label.strata = TRUE) +
  theme_economist() +
  ggtitle("Region and age group by gender", "Equality is everywhere") 
```

# Loan Exploration
Here we investigate if there is any association between the region and the likelihood of default in the 682 loan observations in the dataset.

```{r Loan_status_by_region, echo=TRUE, out.width = '100%'}
left_join(loan, disposition, by = 'account_id') %>% 
  left_join(client, by = 'client_id') %>% 
  left_join(district, by = 'district_id') %>% 
  group_by(region, contract_status, defaulter) %>% 
  summarise(count = n(),
            amount = sum(amount)) %>% 
  group_by(region, contract_status) %>% 
  mutate(count_contract_status = sum(count),
         amount_contract_status = sum(amount)) %>% 
  group_by(region) %>% 
  mutate(count_region = sum(count),
         amount_region = sum(amount)) %>% 
  ggplot(aes(x = defaulter, y = contract_status, fill = count / count_region)) +
  geom_bin2d(stat = 'identity') +
  geom_text(aes(label = paste(round(count / count_region * 100, 2), '%')), 
            color = 'white') +
  facet_wrap(~region) +
  theme_economist() +
  theme(legend.position = 'none', panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(x = 'Defaulter', 
       y = 'Contract Status', 
       title = 'Loan Contract Status by Region Heatmap')
```

We perform the same investigation on the association between the client gender and the likelihood of default in the 682 loan observations in the dataset.

```{r Loan_status_by_gender, echo=TRUE, out.width = '100%'}
left_join(loan, disposition, by = 'account_id') %>%
  left_join(client, by = 'client_id') %>% 
  left_join(district, by = 'district_id') %>% 
  group_by(gender, contract_status, defaulter) %>% 
  summarise(count = n(),
            amount = sum(amount)) %>% 
  group_by(gender, contract_status) %>% 
  mutate(count_contract_status = sum(count),
         amount_contract_status = sum(amount)) %>% 
  group_by(gender) %>% 
  mutate(count_gender = sum(count),
         amount_gender = sum(amount)) %>% 
  ggplot(aes(x = defaulter, y = contract_status, 
             fill = count / count_gender)) +
    geom_bin2d(stat = 'identity') +
    geom_text(aes(label = paste(round(count / count_gender * 100, 2), '%')), 
              color = 'white') +
    facet_wrap(~gender) +
    theme_economist() +
    theme(legend.position = 'none', panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(x = 'Defaulter',
         y = 'Contract Status',
         title = 'Loan Contract Status by Gender Heatmap')
```

We finnaly do the same investigation on the association between the client age and the likelihood of default in the 682 loan observations in the dataset.

```{r Loan_status_by_age_bin, echo=TRUE, out.width = '100%'}
left_join(loan, disposition, by = 'account_id') %>%
  left_join(client, by = 'client_id') %>% 
  left_join(district, by = 'district_id') %>% 
  group_by(age_bin, contract_status, defaulter) %>% 
  summarise(count = n(),
            amount = sum(amount)) %>% 
  group_by(age_bin, contract_status) %>% 
  mutate(count_contract_status = sum(count),
         amount_contract_status = sum(amount)) %>% 
  group_by(age_bin) %>% 
  mutate(count_age_bin = sum(count),
         amount_age_bin = sum(amount)) %>% 
  ggplot(aes(x = defaulter, 
             y = contract_status, fill = count / count_age_bin)) +
    geom_bin2d(stat = 'identity') +
    geom_text(aes(label = paste(round(count / count_age_bin * 100, 2), '%')), 
              color = 'white') +
    facet_wrap(~age_bin) +
    theme_economist() +
    theme(legend.position = 'none', panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(x = 'Defaulter',
         y = 'Contract Status',
         title = 'Loan Contract Status by Age Group Heatmap')
```

It seems that none of these features, alone, are determinant on the odds of a client default.

# Account Balance Exploration

As we do not have the individual income of each client, we tried to perform a proxy of the client's wealth by calculating the average account balance of each account and investigating the distribution by region and age.

```{r avg_account_balance_by_gender_and_region, echo=TRUE, out.width = '100%'}

left_join(account_balance, disposition, by = 'account_id') %>%
  left_join(client, by = 'client_id') %>% 
  left_join(district, by = 'district_id') %>% 
  filter(type == 'Owner') %>% 
  ggplot(aes(avg_balance / 1000)) +
    geom_density(alpha = 0.5, aes(fill = gender)) +
    scale_x_continuous(labels = scales::comma) +
    labs(title = 'Average Account Balance Distribution by Gender and Region',
         subtitle = 'k (Kč) - Koruna (CZK)') +
    theme_economist() +
    facet_wrap(~region)
```

One interest pattern shows up, males in prague region tend to have a bigger average account balance them females, and the oposite trend in South, North and West Bohemia.

# District exploration

Regarding the Czech regions, we can notice in the exploration below which regions are more likely to have defaulters based on the finished loans operations so far.

```{r loan_district_map_exploration, echo=TRUE, out.width = '100%', message=FALSE, warning=FALSE}
loan_amount_by_region <- dplyr::select(loan, account_id, amount, defaulter, contract_status) %>% 
  filter(defaulter == TRUE) %>% 
  inner_join(account) %>% 
  inner_join(district) %>% 
  group_by(region) %>% 
  summarise(transaction_count = n(),
            amount = sum(amount)) %>% 
  inner_join(czech_regions_coords)

jsonMapFile <- "./map/czech-republic-regions.json"
czech_regions <- as.json(geojson_read(jsonMapFile))

leaflet(loan_amount_by_region) %>% 
  addTiles() %>% 
  setView(lng = 15.3, lat = 49.8, zoom = 7) %>%
  addGeoJSON(czech_regions, fillColor = "red", stroke = "#555555") %>% 
  addCircles(lng = ~long, 	
             lat = ~lat, 	
             weight = 2, 	
             radius = ~sqrt(amount) * 30, 	
             popup = ~region)
```
